---
title: Random Forest Analysis
author: Bhawna G. Panwar
date: '2017-07-02'
slug: random-forest-analysis
categories: []
tags: []
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---
The Random Forest analysis techniques was used to improve prediction on the German Credit Dataset.

The dataset is found:
```{r}
url="http://freakonometrics.free.fr/german_credit.csv"
credit=read.csv(url, header = TRUE, sep = ",")
```

```{r}
str(credit)
```
convert categorical variables as factors,
```{r}
F=c(1,2,4,5,7,8,9,10,11,12,13,15,16,17,18,19,20)
for(i in F) credit[,i]=as.factor(credit[,i])
```

create our training/calibration and validation/testing datasets, with proportion 1/3-2/3
```{r}
i_test=sample(1:nrow(credit),size=333)
i_calibration=(1:nrow(credit))[-i_test]
```

```{r}
LogisticModel <- glm(Creditability ~ Account.Balance + Payment.Status.of.Previous.Credit + Purpose + 
Length.of.current.employment + 
Sex...Marital.Status, family=binomial, 
data = credit[i_calibration,])
```

```{r}
fitLog <- predict(LogisticModel,type="response", newdata=credit[i_test,])
```

```{r}
library(ROCR)
pred = prediction( fitLog, credit$Creditability[i_test])
perf <- performance(pred, "tpr", "fpr")
plot(perf)
AUCLog1=performance(pred, measure = "auc")@y.values[[1]]
cat("AUC: ",AUCLog1,"n")
```

Alternative is to consider a logistic regression on all explanatory variables
```{r}
LogisticModel <- glm(Creditability ~ ., family=binomial, data = credit[i_calibration,])
fitLog <- predict(LogisticModel,type="response",newdata=credit[i_test,])
pred = prediction( fitLog, credit$Creditability[i_test])
perf <- performance(pred, "tpr", "fpr")
plot(perf)
AUCLog2=performance(pred, measure = "auc")@y.values[[1]]
cat("AUC: ",AUCLog2,"n")
```

Regression tree method
```{r}
library(rpart)
ArbreModel <- rpart(Creditability ~ ., data = credit[i_calibration,])
```
```{r}
library(rpart.plot)
prp(ArbreModel,type=2,extra=1)
```
```{r}
fitArbre <- predict(ArbreModel, newdata=credit[i_test,],type="prob")[,2]
pred = prediction( fitArbre, credit$Creditability[i_test])
perf <- performance(pred, "tpr", "fpr")
plot(perf)
AUCArbre=performance(pred, measure = "auc")@y.values[[1]]
cat("AUC: ",AUCArbre,"n")
```
Random Forest analysis
```{r}
library(randomForest)
RF <- randomForest(Creditability ~ .,data = credit[i_calibration,])
fitForet <- predict(RF,newdata=credit[i_test,],type="prob")[,2]
pred = prediction( fitForet, credit$Creditability[i_test])
perf <- performance(pred, "tpr", "fpr")
plot(perf)
AUCRF=performance(pred, measure = "auc")@y.values[[1]]
cat("AUC: ",AUCRF,"n")
```