---
title: Association Rules
author: ~
date: '2017-07-02'
slug: association-rules-rmd
categories: []
tags: []
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---

###############################################

ASSOCIATION RULES
  
###############################################
Association Rules-identifying frequently purchased groceries with association rules.
market basket analysis will utilize the purchase data collected from one month of operation at a real-world grocery store. The data contains 9,835 transactions or about 327 transactions per day (roughly 30 transactions per hour in a 12-hour business day), suggesting that the retailer is not particularly large, nor is it particularly small.

The typical grocery store offers a huge variety of items. There might be five brands of milk, a dozen different types of laundry detergent, and three brands of coffee. Given the moderate size of the retailer, we will assume that they are not terribly concerned with finding rules that apply only to a specific brand of milk or detergent. With this in mind, all brand names can be removed from the purchases. This reduces the number of groceries to a more manageable 169 types, using broad categories such as chicken, frozen meals, margarine, and soda.

Since we're loading the transactional data, we cannot simply use the read.csv() function used previously. Instead, arules provides a read.transactions() function that is similar to read.csv() with the exception that it results in a sparse matrix suitable for transactional data. The sep = "," parameter specifies that items in the input file are separated by a comma. To read the groceries.csv data into a sparse matrix named groceries, type the following line:


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
##### Chapter 8: Association Rules 

## Example: Identifying Frequently-Purchased Groceries 
## Step 2: Exploring and preparing the data 

# load the grocery data into a sparse matrix
library(arules)
groceries <- read.transactions("C:/Users/Bhawna/Documents/blog/data/groceries.csv", sep = ",")
summary(groceries)

# look at the first five transactions
inspect(groceries[1:5])

# examine the frequency of items
itemFrequency(groceries[, 1:3])

# plot the frequency of items
itemFrequencyPlot(groceries, support = 0.1)
itemFrequencyPlot(groceries, topN = 20)

# a visualization of the sparse matrix for the first five transactions
image(groceries[1:5])

# visualization of a random sample of 100 transactions
image(sample(groceries, 100))

```

```{r}
## Step 3: Training a model on the data 
library(arules)

# default settings result in zero rules learned
apriori(groceries)

# set better support and confidence levels to learn more rules
groceryrules <- apriori(groceries, parameter = list(support =
                          0.006, confidence = 0.25, minlen = 2))
groceryrules

```

```{r}
## Step 4: Evaluating model performance 
# summary of grocery association rules
summary(groceryrules)

# look at the first three rules
inspect(groceryrules[1:3])
```

```{r}
## Step 5: Improving model performance 

# sorting grocery rules by lift
inspect(sort(groceryrules, by = "lift")[1:5])

# finding subsets of rules containing any berry items
berryrules <- subset(groceryrules, items %in% "berries")
inspect(berryrules)
```

Summary

Association rules are frequently used to find provide useful insights in the massive transaction databases of large retailers. The Apriori algorithm, which we studied in this chapter, does so by setting minimum thresholds of uniqueness, and reporting only the associations meeting these criteria.

Reference: Machine Learning with R