---
title: Clustering Analysis
author: ~
date: '2017-07-02'
slug: clustering-analysis
categories: []
tags: []
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
---


###############################################

CLUSTERING ANALYSIS
  
###############################################

The cluster Analysis is performed on the SNS dataset.Clustering with k-means can be applied by identifying groups. Clustering can automate the process of discovering the natural segments in this population. However, it will be up to us to decide whether or not the clusters are interesting and how we can use them for advertising.

SNS dataset representing a random sample of 30,000 U.S. high school students who had profiles on a well-known SNS in 2006. To protect the users' anonymity, the SNS will remain unnamed. However, at the time the data was collected, the SNS was a popular web destination for US teenagers. Therefore, it is reasonable to assume that the profiles represent a fairly wide cross section of American adolescents in 2006.

The data was sampled evenly across four high school graduation years (2006 through 2009) representing the senior, junior, sophomore, and freshman classes at the time of data collection. Using an automated web crawler, the full text of the SNS profiles were downloaded, and each teen's gender, age, and number of SNS friends was recorded.

A text mining tool was used to divide the remaining SNS page content into words. From the top 500 words appearing across all the pages, 36 words were chosen to represent five categories of interests: namely extracurricular activities, fashion, religion, romance, and antisocial behavior. The 36 words include terms such as football, sexy, kissed, bible, shopping, death, and drugs. The final dataset indicates, for each person, how many times each word appeared in the person's SNS profile.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
##### Chapter 9: Clustering with k-means 

## Example: Finding Teen Market Segments ----
## Step 2: Exploring and preparing the data ----
teens <- read.csv("C:/Users/Bhawna/Documents/blog/data/snsdata.csv")
str(teens)

# look at missing data for female variable
table(teens$gender)
table(teens$gender, useNA = "ifany")

# look at missing data for age variable
summary(teens$age)

# eliminate age outliers
teens$age <- ifelse(teens$age >= 13 & teens$age < 20,
                     teens$age, NA)

summary(teens$age)

# reassign missing gender values to "unknown"
teens$female <- ifelse(teens$gender == "F" &
                         !is.na(teens$gender), 1, 0)
teens$no_gender <- ifelse(is.na(teens$gender), 1, 0)

# check our recoding work
table(teens$gender, useNA = "ifany")
table(teens$female, useNA = "ifany")
table(teens$no_gender, useNA = "ifany")

# finding the mean age by cohort
mean(teens$age) # doesn't work
mean(teens$age, na.rm = TRUE) # works

# age by cohort
aggregate(data = teens, age ~ gradyear, mean, na.rm = TRUE)

# create a vector with the average age for each gradyear, repeated by person
ave_age <- ave(teens$age, teens$gradyear,
                 FUN = function(x) mean(x, na.rm = TRUE))


teens$age <- ifelse(is.na(teens$age), ave_age, teens$age)

# check the summary results to ensure missing values are eliminated
summary(teens$age)
```

```{r}
## Step 3: Training a model on the data ----
interests <- teens[5:40]
interests_z <- as.data.frame(lapply(interests, scale))

set.seed(2345)
teen_clusters <- kmeans(interests_z, 5)

```
```{r}
## Step 4: Evaluating model performance ----
# look at the size of the clusters
teen_clusters$size

# look at the cluster centers
teen_clusters$centers


```
```{r}
## Step 5: Improving model performance ----
# apply the cluster IDs to the original data frame
teens$cluster <- teen_clusters$cluster

# look at the first five records
teens[1:5, c("cluster", "gender", "age", "friends")]

# mean age by cluster
aggregate(data = teens, age ~ cluster, mean)

# proportion of females by cluster
aggregate(data = teens, female ~ cluster, mean)

# mean number of friends by cluster
aggregate(data = teens, friends ~ cluster, mean)
```
Reference: Machine Learning with R - Second Edition - Jul 31, 2015 by Brett Lantz